### 2021-03-02 conclusions.tex
Conclusions.

\section{\label{sec:concl}Conclusions and Future Work}

In this paper we investigate the notion of extraction.  This
is the ability to express programs hand-in-hand with custom
code generators for them.  We have demonstrated that
compile-time metaprogramming is sufficient to equip
a programming language with extraction capabilities.
A well-known conundrum of choosing between deep and
shallow embedding can be resolved if the host language
is extraction-capable.  While enjoying the benefits
of shallow embedding, there is a full access to the internal
structure of the embedded programs.

We use this idea in the context of dependently-typed
languages to create verified implementations that can
be used in the context of an existing programming language.
We embed a language of interest in a theorem-prover, using
dependent types to encode the properties of interest.
Using extraction we bring the verified implementation
back into the original language.

We have demonstrated the approach by implementing
three embedded languages and two extractors using
Agda as our host.  We have demonstrated the
changes to Agda that we had to do on the way, and we
used our embedding to implement (and verify) a practical
application --- a convolutional neural network.

The main advantages are twofold.  First,
our solution is fine-grain --- we can chose what part
of the application to embed, and what constructs of the
host language to extract.  Secondly, extractors
can use the full power of dependent types guaranteeing
some safety properties of interest.

In order to make the proposed idea even more powerful,
we would require improvements to metaprogramming capabilities
of theorem provers.

Most importantly, right now it is not guaranteed that the
extracted code preservers the semantics of the original
implementation.  Essentially, one has to trust that the
extractor is doing the right job.  While we rarely see
fully-verified compiler backends in the real world, in
the proposed environment we are very close to enabling this.
We would need a formal semantics of the reflected language
and the proof that reflected programs respect it.
While this is non-trivial, a system like Agda could do
this in principle.

There is a number of improvements that can be added to
Agda and our extractors to make the resulting code more
efficient.  Supporting `let's in the internal syntax
would help to preserve sharing.  Recognising irrelevance
annotations in the extractors would help to eliminate
unused function arguments.  Introducing proper language
primitives to specify what exactly is an embedding would
be helpful.

Overall, this work only scratches the surface of extraction-based
compilation.  We never considered alternative theories supported
by theorem provers, \eg{} cubical type theory in Agda; we did not
consider recursive metaprogramming; we did not consider integrating
optimisations of extracted programs besides of what rewriting rules
are capable to do.  All of these offer exciting research opportunities
on the way to make verified software easily accessible.
To paraphrase Jim Morrison: ``Safety and no surprise, The End''.